---
title: "EID Updates"
author: "Daniel Fedha"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  word_document:
    highlight: tango
    toc: yes

---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, include = T, warning = F, message = F, comment = NA, collapse = T)
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "C:/Program Files/MiKTeX 2.9/miktex/bin/x64", sep=.Platform$path.sep))
```

```{r loadpackages}
library(ggplot2)
library(readr)
library(dplyr)
library(writexl)
library(knitr)
library(flextable)
library(tidyr)
library(lubridate)
```

######
######
######
######
######
&nbsp;



```{r sqlstatements, echo=FALSE, include=FALSE}
#todate <- Sys.time()
# 
#  if(!require('RPostgreSQL')) install.packages("RPostgreSQL", dependencies = T)
#    require("RPostgreSQL")
#
#    drv <- dbDriver("PostgreSQL")
#
#    con <- dbConnect(drv, dbname = "okoa",
#                    host = "localhost", port = 5432,
#                     user = "okoauser", password ='0k0@us3r') #rstudioapi::askForPassword("Database password"))
#    on.exit(dbDisconnect(con), add = TRUE)
#
#    #check for a table dbExistsTable(con, "okoa_laboratorysample")
#
#    sqlQuery = "SELECT	ls.study_identification_number as sid,
#  CONCAT(a_user.first_name, ' ', a_user.last_name) as staff_name,
#  lab.name as lab_name,   ls.dbs_collection_date as date_dbscollected,   ls.viable_blood_spots_count as viable_bloodspots,
#  ls.date_of_birth as infant_dob,   ls.date_sample_details_added as date_sampleadded,   ls.infant_gender,
#  sep.name as sample_entrypoint,   ls.sample_entry_point_other as sample_entrypoint_other,
#  ipr.name as infantprophylaxis,   ifi.name as infant_feedinginfo,
#  ls.infant_prophylaxis_regimen_other as infantprophylaxis_other,
#  DATE_PART('day',dbs_collection_date::timestamp - date_of_birth::timestamp) as infantage_days,
#  ls.mother_age,   ls.mother_alive_status as mother_status,   ls.maternal_haart_initiation_date as motherhaartinitiation_date,
#  pmtct.name as  mother_pmtctregimen,   ls.maternal_pmtct_regimen_other as mother_pmtctregimen_other,
#  pcr.name as pcrtest_requested,   ls.sample_redraw_pcr_code_id as sampleredraw_pcr,
#  ls.user_receiving_sample_dr_lab_id as staff_receivingsample,   ls.dr_lab_no_viable_blood_spots  as viablebloodspots_dr,
#  ls.date_genotyping_dbs_spots as dbsgenotyping_date,   ls.pcr_test_positive_result_date as pcrtestpositiveresult_date,
#  ls.dbs_placed_in_freezer_date as dbsplacedinfreezer_date,   ls.success_sequencing_rt_pol1_region,
#  ls.success_sequencing_pr_pol1_region,   rsu.name as reasonsample_unacceptable,   rpc.name as receiptpackaging_condition,
#  sm.shipping_date,   sm.receiving_date,   sm.receiving_time,   sm.date_placed_in_freezer as dateplaced_freezer,
#  sm.time_placed_in_freezer as timeplaced_freezer,   sm.shipping_temp,   sm.receiving_temp,   sms.name as manifest_status
#
#    FROM
#    public.okoa_laboratorysample as ls
#
#    left join okoa_motherpmtctregimen as pmtct
#    on ls.maternal_pmtct_regimen_id = pmtct.id
#
#    left join okoa_sampleentrypoint as sep
#    on ls.sample_entry_point_id = sep.id
#
#    left join okoa_infantfeedinginformation as ifi
#    on ls.infant_feeding_information_id = ifi.id
#
#    left join okoa_infantprophylaxisregimen as ipr
#    on ls.infant_prophylaxis_regimen_id = ipr.id
#
#    left join okoa_pcrtest as pcr
#    on ls.requested_pcr_test_id = pcr.id
#
#    left join okoa_laboratory as lab
#    on ls.laboratory_id = lab.id
#
#    left join okoa_samplemanifestline as sml
#    on ls.id = sml.laboratory_sample_id
#
#    left join okoa_samplemanifest as sm
#    on sm.id = sml.sample_manifest_id
#
#    left join okoa_samplemanifeststatus as sms
#    on sms.id = sm.manifest_status_id
#
#    left join okoa_reasonsampleunacceptable as rsu
#    on rsu.id = sml.reason_sample_unacceptable_id
#
#    left join okoa_receiptpackagingconditioncode as rpc
#    on rpc.id = sml.receipt_packaging_condition_id
#
#    left join auth_user as a_user
#   on a_user.id = ls.user_entering_sample_details_id;"
#
#
#     data <- dbGetQuery(con, sqlQuery)
#
#
#      #close db connection
#      #lapply(dbListConnections(drv = dbDriver("PostgreSQL")), function(x) {dbDisconnect(conn = x)})
#
#     data$enrolled = 1
#     saveRDS(data, file='C:/Users/Daniel Fedha/Box Sync/data/EID/data.rds')


```


```{r data-processing, echo=FALSE}

data <- readRDS('C:/Users/Daniel Fedha/Box Sync/data/EID/data.rds')
data$date_dbscollected <- as.Date(data$date_dbscollected, format = '%d/%m/%Y')
data$date_sampleadded <- as.Date(data$date_sampleadded, format = '%d/%m/%Y')
data$dbsplacedinfreezer_date<- as.Date(data$dbsplacedinfreezer_date, format = '%d/%m/%Y')
data$dbsgenotyping_date <- as.Date(data$dbsgenotyping_date, format = '%d/%m/%Y')
data$pcrtestpositiveresult_date <- as.Date(data$pcrtestpositiveresult_date, format = '%d/%m/%Y')
data$infant_dob <- as.Date(data$infant_dob, format = '%d/%m/%Y')
data$lab_name <- as.factor(data$lab_name)
data$infant_gender <- as.factor(data$infant_gender)
data$sample_entrypoint <- as.factor(data$sample_entrypoint)
data$infantprophylaxis <- as.factor(data$infantprophylaxis)
data$infant_feedinginfo <- as.factor(data$infant_feedinginfo)
data$mother_status <- as.factor(data$mother_status)
data$mother_pmtctregimen <- as.factor(data$mother_pmtctregimen)
data$pcrtest_requested <- as.factor(data$pcrtest_requested)
data$week_enrol <- 0
data$week_enrol[data$date_sampleadded >= '2018/06/01' & data$date_sampleadded <= '2018/06/09'] <- 1
data$week_enrol[data$date_sampleadded >= '2018/06/10' & data$date_sampleadded <= '2018/06/16'] <- 2
data$week_enrol[data$date_sampleadded >= '2018/06/17' & data$date_sampleadded <= '2018/06/23'] <- 3
data$week_enrol[data$date_sampleadded >= '2018/06/24' & data$date_sampleadded <= '2018/06/30'] <- 4
data$week_enrol[data$date_sampleadded >= '2018/07/01' & data$date_sampleadded <= '2018/07/07'] <- 5
data$week_enrol[data$date_sampleadded >= '2018/07/08' & data$date_sampleadded <= '2018/07/14'] <- 6
data$week_enrol[data$date_sampleadded >= '2018/07/15' & data$date_sampleadded <= '2018/07/21'] <- 7
data$week_enrol[data$date_sampleadded >= '2018/07/22' & data$date_sampleadded <= '2018/07/28'] <- 8
data$week_enrol[data$date_sampleadded >= '2018/07/29' & data$date_sampleadded <= '2018/08/04'] <- 9
data$week_enrol[data$date_sampleadded >= '2018/08/05' & data$date_sampleadded <= '2018/08/11'] <- 10
data$week_enrol[data$date_sampleadded >= '2018/08/12' & data$date_sampleadded <= '2018/08/18'] <- 11
data$week_enrol[data$date_sampleadded >= '2018/08/19' & data$date_sampleadded <= '2018/08/25'] <- 12
data$week_enrol[data$date_sampleadded >= '2018/08/26' & data$date_sampleadded <= '2018/09/01'] <- 13
data$week_enrol[data$date_sampleadded >= '2018/09/02' & data$date_sampleadded <= '2018/08/08'] <- 14
data$week_enrol[data$date_sampleadded >= '2018/09/09' & data$date_sampleadded <= '2018/09/15'] <- 15
data$week_enrol[data$date_sampleadded >= '2018/09/16' & data$date_sampleadded <= '2018/09/22'] <- 16
data$week_enrol[data$date_sampleadded >= '2018/09/23' & data$date_sampleadded <= '2018/09/30'] <- 17
data$week_enrol[data$date_sampleadded >= '2018/10/01' & data$date_sampleadded <= '2018/10/06'] <- 18
data$week_enrol[data$date_sampleadded >= '2018/10/07' & data$date_sampleadded <= '2018/10/13'] <- 19
data$week_enrol[data$date_sampleadded >= '2018/10/14' & data$date_sampleadded <= '2018/10/20'] <- 20
data$week_enrol[data$date_sampleadded >= '2018/10/21' & data$date_sampleadded <= '2018/10/27'] <- 21
data$week_enrol[data$date_sampleadded >= '2018/10/28' & data$date_sampleadded <= '2018/10/31'] <- 22
data$week_enrol[data$date_sampleadded >= '2018/11/01' & data$date_sampleadded <= '2018/11/10'] <- 23
data$week_enrol[data$date_sampleadded >= '2018/11/11' & data$date_sampleadded <= '2018/11/17'] <- 24
data$week_enrol[data$date_sampleadded >= '2018/11/18' & data$date_sampleadded <= '2018/11/24'] <- 25
data$week_enrol[data$date_sampleadded >= '2018/11/25' & data$date_sampleadded <= '2018/11/30'] <- 26
data$week_enrol[data$date_sampleadded >= '2018/12/01' & data$date_sampleadded <= '2018/12/08'] <- 27
data$week_enrol[data$date_sampleadded >= '2018/12/09' & data$date_sampleadded <= '2018/12/15'] <- 28
data$week_enrol[data$date_sampleadded >= '2018/12/16' & data$date_sampleadded <= '2018/12/22'] <- 29
data$week_enrol[data$date_sampleadded >= '2018/12/23' & data$date_sampleadded <= '2018/12/29'] <- 30

my_theme <- function(){
  theme_light() +
    theme(panel.grid = element_blank(),
          #panel.border = element_blank(),
          # axis.title = element_blank(),
          # axis.ticks = element_blank(),
          # axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_text(size = 8, colour = 'black'))
}

sep_index <- data$sample_entrypoint == 'Other'
pmt_index <- data$mother_pmtctregimen == 'Other'
ipr_index <- data$infantprophylaxis == 'Other'

data$infantprophylaxis <- as.character(data$infantprophylaxis)
data$sample_entrypoint <- as.character(data$sample_entrypoint)
data$mother_pmtctregimen <- as.character(data$mother_pmtctregimen)


data$infantprophylaxis[ipr_index] <- paste('Other - ', data$infantprophylaxis_other[ipr_index])
data$sample_entrypoint[sep_index] <- paste('Other - ',data$sample_entrypoint_other[sep_index])
data$sample_entrypoint[data$sample_entrypoint_other=='IPD'] <- 'Paediatric Ward/ IPD'
data$mother_pmtctregimen[pmt_index] <- paste('Other - ', data$mother_pmtctregimen_other[pmt_index])
data$mother_pmtctregimen[data$mother_pmtctregimen_other=='Not provided'] <- 'Unknown'

rm(sep_index, pmt_index, ipr_index)

```

```{r data-cleaning}
data$sid[data$sid=='01EIDKE180018'] <- '01EIDKE18018'
data$sid[data$sid=='02EIDKE180014'] <- '02EIDKE18014'

```

```{r supervision}

# #  set.seed(10) #for reproducibility
# #  df2 <- df[sample(nrow(df), 3), ] #Select 3 random rows
# # sample_n((0.1*nrow(alupedt)))
# 
# supData <- data %>% 
#   select(sid, date_sampleadded, staff_name, lab_name,
#          date_dbscollected:sample_entrypoint, infantprophylaxis,
#          infant_feedinginfo, 
#          mother_age:mother_pmtctregimen, pcrtest_requested, 
#          pcrtestpositiveresult_date, dateplaced_freezer)
# 
# #-------------------
# coastdt <- supData %>% filter(lab_name=='Coast Provincial General Hospital')
# alupedt <- supData%>% filter(lab_name=='KEMRI CIPDCR, Alupe Busia') 
# kisiandt <- supData %>% filter(lab_name=='KEMRI CDC HIV/R, Kisumu') 
# ampathdt <- supData %>% filter(lab_name=='AMPATH, Eldoret') 
# nhrldt <- supData %>% filter(lab_name=='National HIV Reference Laboratory') 
# p3dt <- supData %>% filter(lab_name=='KEMRI CVR HIV-P3, Nairobi') 
# 
# write_xlsx (list('CPGH'=coastdt, 'ALUPE'=alupedt, 'KISIAN'=kisiandt, 'AMPATH'=ampathdt, 'P3'=p3dt, 'NHRL'=nhrldt), 'C:/Users/Daniel Fedha/Box Sync/data/EID/supervisionwb.xlsx')
# 
# rm(supData, coastdt, alupedt, p3dt, nhrldt, ampathdt, kisiandt)
```

# 1. **Enrolment Statistics**

a. ## Enrolment by Lab - Table:

```{r enrolLab, fig.height=2, fig.width=6}

sampleSize <- read_csv(file = 'C:/Users/Daniel Fedha/Box Sync/data/EID/sample_allocation.csv')

dtm <- merge(data %>%
               group_by(lab_name) %>%
                select(lab_name, enrolled, week_enrol) %>%
                filter(week_enrol<=4) %>%
                summarise('June'=sum(enrolled)),
             data %>%
                    group_by(lab_name) %>%
                    select(lab_name, enrolled, week_enrol) %>% 
                    filter(week_enrol>=5 & week_enrol <=8) %>%
                    summarise('July'=sum(enrolled)),
        by='lab_name', all = T) %>%
  
  merge(data %>%
            group_by(lab_name) %>%
            select(lab_name, enrolled, week_enrol) %>% 
            filter(week_enrol >=9 & week_enrol <=13) %>%
            summarise('Aug'=sum(enrolled)),
        by='lab_name', all = T) %>%
  
  merge(data %>%
            group_by(lab_name) %>%
            select(lab_name, enrolled, week_enrol) %>%
            filter(week_enrol>=14 & week_enrol <=17) %>%
            summarise('Sep'=sum(enrolled)),
        by='lab_name', all = T) %>%
  
  merge(data %>%
            group_by(lab_name) %>%
            select(lab_name, enrolled, week_enrol) %>%
            filter(week_enrol>=18 & week_enrol<=22) %>%
            summarise('Oct'=sum(enrolled)),
        by='lab_name', all = T) %>%
   
   merge(data %>%
            group_by(lab_name) %>%
            select(lab_name, enrolled, week_enrol) %>%
            filter(week_enrol>=23 & week_enrol<=26) %>%
            summarise('Nov'=sum(enrolled)),
        by='lab_name', all = T) %>%
  
   merge(data %>%
          group_by(lab_name) %>%
          select(lab_name, enrolled, week_enrol) %>%
          filter(week_enrol>26) %>%
          summarise('Dec'=sum(enrolled)),
      by='lab_name', all = T) %>%
  
  
  replace_na(list('June'=0, 'July'=0, 'Aug'=0, 'Sep'=0, 'Oct'=0, 'Nov'=0, 'Dec'=0))

dtm <- merge(dtm, sampleSize, by='lab_name') %>%  rename('Lab Name'=lab_name)

dtm$allocated_sampleSize <- dtm$allocated_sampleSize
dtm$Total <- dtm$`June`+ dtm$`July`+ dtm$`Aug`+ dtm$`Sep`+dtm$`Oct`+dtm$`Nov`+dtm$`Dec`
dtm$difference <- dtm$Total - dtm$allocated_sampleSize
dtm$percentage_achieved <- (dtm$Total/dtm$allocated_sampleSize)*100
dtm$percentage_achieved <- round(dtm$percentage_achieved) 
dtm <- dtm %>% arrange(desc(percentage_achieved))
dtm$`Lab Name` <- as.character(dtm$`Lab Name`)
dtm <- rbind(dtm, c("Total", colSums(dtm[,-c(1,12)]), 0))
dtm$Total <- as.integer(dtm$Total)
dtm$percentage_achieved[dtm$allocated_sampleSize==560] <-round((sum(dtm$Total[1:8])/560)*100)

dtm %>% select('Lab Name', `June`, `July`, `Aug`, `Sep`, `Oct`, `Nov`, `Dec`, Total, 
               allocated_sampleSize, difference, percentage_achieved) %>%
  rename('% Achieved'= percentage_achieved, 'Sample Size'=allocated_sampleSize) %>%
  regulartable() %>%  theme_zebra() %>%  autofit()

```
Summary:&nbsp;
```{r summary}
sampEP <- aggregate(data$enrolled, by=list(sEP=data$sample_entrypoint), FUN=sum)
infaF <- aggregate(data$enrolled, by=list(infaF=data$infant_feedinginfo), FUN=sum)
pcrT <- aggregate(data$enrolled, by=list(pcrT=data$pcrtest_requested), FUN=sum)
pmtct <- aggregate(data$enrolled, by=list(infaF=data$mother_pmtctregimen), FUN=sum)

```


* `r sum(as.integer(dtm$Total[1:8]))` samples have been enrolled over a period of 6 months &nbsp;
  * `r subset(dtm, dtm$Total==max(as.integer(dtm[,'Total'][1:8])), select='Lab Name')` has enrolled the highest no. of samples: `r max(as.integer(dtm[,'Total'][1:8]))`, &nbsp;

  * most samples [N=`r max(sampEP[,'x']) `, 
  (`r format(((max(sampEP[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)`%)] are through MCH/PMTCT sample entry point &nbsp;
  
  * most samples [N= `r max(infaF[,'x']) `,
  (`r format(((max(infaF[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)`%] are on EBF &nbsp;
  
  * Initial PCR at 6 weeks or first contact is the most requested [N= `r max(pcrT[,'x']) `, 
  (`r format(((max(pcrT[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)`%] &nbsp;
  
  * most mothers [N= `r max(pmtct[,'x']) `, (`r format(((max(pmtct[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)` %] are on TDF+3TC+EFV &nbsp;

  
  * Missing data for primary variables is 0% &nbsp;

&nbsp;

b. ##  Lab Enrolment by Gender



```{r enrolBar, fig.height=3, fig.width=8}
data %>%
    group_by(lab_name, infant_gender) %>%
    select(lab_name, enrolled, infant_gender) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(lab_name, enrolled),
             y=enrolled, fill=infant_gender, position='stack')) +
  geom_bar(width=0.8, stat = 'identity') +
  geom_text(aes(label = enrolled, y=enrolled), position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0)) +  # Set the axes to cross at 0
  labs(x='Laboratory', y='No. Enrolled') +
  my_theme() 

```

 &nbsp;
 
 c. ## Sample Entry Point by Lab:

```{r enrolBSampEntryP, fig.height=4, fig.width=12}
data %>%
    group_by(sample_entrypoint, lab_name) %>%
    select(lab_name, enrolled, sample_entrypoint) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(sample_entrypoint, enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 300, 20)) +  # Set the axes to cross at 0
  
  labs(x='Laboratory',
       y=' ') +
  my_theme() 

```
&nbsp;

d. ## Infant Feeding by Lab:

```{r fig.height=3, fig.width=8}
data %>%
    group_by(infant_feedinginfo, lab_name) %>%
    select(lab_name, enrolled, infant_feedinginfo) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(infant_feedinginfo,  enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 250, 20)) +  # Set the axes to cross at 0
  labs(x='Laboratory',
       y=' ') +
  my_theme() 

```

&nbsp;

  e. ## PCR Test Requested by Lab:

```{r fig.height=3, fig.width=12}
data %>%
    group_by(pcrtest_requested, lab_name) %>%
    select(lab_name, enrolled, pcrtest_requested) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(pcrtest_requested, enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 400, 30)) +  # Set the axes to cross at 0
  labs(x='Laboratory',
       y=' ') +
  my_theme() 

```
&nbsp;

  f. ## Mother PMTCT Regimen by Lab:
```{r fig.height=6, fig.width=12}

data %>%
    group_by(mother_pmtctregimen, lab_name) %>%
    select(lab_name, enrolled, mother_pmtctregimen) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(mother_pmtctregimen, enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.9, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 340, 20)) +  # Set the axes to cross at 0
  labs(x='Laboratory',
       y=' ') +
  my_theme() 


```
&nbsp;

g. ## Age distribution - Histogram [bin-width=30 days]

```{r enrolHist, fig.height=3, fig.width=6}

ggplot(data, 
       aes(data$infantage_days)) +
  geom_histogram(breaks=seq(1, 547, by =30), col="red", aes(fill=..count..)) +
  scale_fill_gradient("Count", low = "green", high = "red") +
  labs(x='Age in Days', y='Count') +
  scale_y_continuous(expand = c(0, 0)) +  # Set the axes to cross at 0
  scale_x_continuous(expand = c(0, 0)) +
  theme_light() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_text(size = 8, colour = 'black'))
```
&nbsp;



# **2. Tracking Log Data:**

```{r shippingDates}
shipDate <- data %>% filter(!is.na(receiving_date)) %>% select(lab_name, enrolled, starts_with('shipping_'), starts_with('receiving_'))
shipDate$datediff <- difftime(shipDate$receiving_date, shipDate$shipping_date, units = 'days')

shipDate <- merge(
  shipDate %>% group_by(lab_name) %>% select(lab_name, datediff, enrolled) %>% filter(datediff ==0) %>% 
    summarise('Dispatch to Receipt = 0 Days'=sum(enrolled)),
  shipDate %>% group_by(lab_name) %>% select(lab_name, datediff, enrolled) %>% filter(datediff ==1) %>% 
    summarise('Dispatch to Receipt = 1 Day'=sum(enrolled)),
by='lab_name', all=T) %>% 
  merge(
    shipDate %>% group_by(lab_name) %>% select(lab_name, datediff, enrolled) %>% filter(datediff ==2) %>%
      summarise('Dispatch to Receipt = 2 Days' = sum(enrolled)),
    by='lab_name', all=T) %>%

  rename('Lab Name'=lab_name) #%>%

```

```{r dispReceive_DR}
drlab <- merge(
    data %>%
    group_by(lab_name) %>%
    select('Lab Name'=lab_name, manifest_status, enrolled) %>%
    filter(is.na(manifest_status) | manifest_status=='Draft at EID Lab') %>%
    summarise('Not yet Dispatched'=sum(enrolled)),
  
  data %>%
    group_by(lab_name) %>%
    select('Lab Name'=lab_name, manifest_status, enrolled) %>%
    filter(manifest_status =='Sent to DR Lab') %>%
    summarise('Dispatched to DR Lab but not yet Recieved'=sum(enrolled)),
  by='Lab Name', all=T) %>%
  
  merge(
  data %>%
    group_by(lab_name) %>%
    select('Lab Name'=lab_name, manifest_status, enrolled) %>%
    filter(manifest_status == 'Received at DR Lab') %>%
    summarise('Dispatched and Received at DR Lab'=sum(enrolled)),
  by='Lab Name', all=T) %>%
  
  merge(
    data %>%
      group_by(lab_name) %>%
      select('Lab Name'=lab_name, receiptpackaging_condition, enrolled) %>%
      filter(receiptpackaging_condition == 'Good') %>%
      summarise("Packaging Condition is Good"=sum(enrolled)),
  by='Lab Name', all=T) %>%

  merge(
    data %>%
      group_by(lab_name) %>%
      select('Lab Name'=lab_name, receiptpackaging_condition, enrolled) %>%
      filter(receiptpackaging_condition == 'Not Good') %>%
      summarise("Packaging Condition is Not Good"=sum(enrolled)),
  by='Lab Name', all=T)%>%

  merge(
    data %>%
      group_by(lab_name) %>% filter(receiptpackaging_condition == 'Not Good') %>%
      select('Lab Name'=lab_name, 'Reason sample Unaccepted'= reasonsample_unacceptable),
  by='Lab Name', all=T)  

drlab <- merge(drlab, shipDate, by='Lab Name', all = T) %>% select(c(1:4, 8:10, 5:7)) %>%
  replace_na(list('Not yet Dispatched'= 0,
                  'Dispatched to DR Lab but not yet Recieved'=0,
                  'Dispatched and Received at DR Lab'=0,
                  'Packaging Condition is Good'=0,
                  'Packaging Condition is Not Good'=0,
                  'Reason sample Unaccepted'= 'N/A',
                  'Dispatch to Receipt = 0 Days'=0, 
                  'Dispatch to Receipt = 1 Day'=0, 
                  'Dispatch to Receipt = 2 Days'=0))

drlab$`Lab Name` <- as.character(drlab$`Lab Name`)
drlab <- rbind(drlab,c('Total No of Samples', colSums(drlab[,c(2:9)]),'-'))
#drlab[,c(2:5)] <- sapply(drlab[,c(2:5)],as.character)
drlab %>% regulartable() %>% theme_zebra() 

```



# **3. Missing Data:**

N=total no. missing per variable, &nbsp;

% = percentage of missing data per variable

```{r missingSID}
 missD <- merge(
   data %>%
     group_by(lab_name) %>%
     select(sid) %>% 
     summarise('sid'=sum(is.na(.))),
   data %>%
     group_by(lab_name) %>%
     select(date_dbscollected) %>% 
     summarise('date_dbscollected'=sum(is.na(.))),
   by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(viable_bloodspots) %>% 
      summarise('viable_bloodspots'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(infant_dob) %>% 
      summarise('infant_dob'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(infantprophylaxis) %>%
      summarise('infantprophylaxis'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(mother_pmtctregimen) %>% 
      summarise('mother_pmtctregimen'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(pcrtest_requested) %>% 
      summarise('pcrtest_requested'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(pcrtestpositiveresult_date) %>%
      summarise('pcrtestpositiveresult_date'=sum(is.na(.))),
    by='lab_name', all=T)

dtm2 <- dtm %>%
  na.omit() %>%
  select('Lab Name', Total) %>%
  rename(lab_name='Lab Name')
missD <- merge(missD, dtm2, by='lab_name', all = T)
rm(dtm2)
missD$sid_perc <- (missD$sid/as.integer(missD$Total))*100
missD$date_dbscollected_perc <- (missD$date_dbscollected/as.integer(missD$Total))*100
missD$viable_bloodspots_perc <- (missD$viable_bloodspots/as.integer(missD$Total))*100
missD$infant_dob_perc <- (missD$infant_dob/as.integer(missD$Total))*100
missD$infantprophylaxis_perc <- (missD$infantprophylaxis/as.integer(missD$Total))*100
missD$mother_pmtctregimen_perc <- (missD$mother_pmtctregimen/as.integer(missD$Total))*100
missD$pcrtest_requested_perc <- (missD$pcrtest_requested/as.integer(missD$Total))*100
missD$pcrtestpositiveresult_date_perc <- (missD$pcrtestpositiveresult_date/as.integer(missD$Total))*100

missD$sid <- as.character(missD$sid)
missD$sid_perc <- as.character(missD$sid_perc)
missD$sid <- paste(missD$sid, '(', missD$sid_perc, '%', ')')

missD$date_dbscollected <- as.character(missD$date_dbscollected)
missD$date_dbscollected_perc <- as.character(missD$date_dbscollected_perc)
missD$date_dbscollected <- paste(missD$date_dbscollected, '(', missD$date_dbscollected_perc, '%', ')')

missD$viable_bloodspots <- as.character(missD$viable_bloodspots)
missD$viable_bloodspots_perc <- as.character(missD$viable_bloodspots_perc)
missD$viable_bloodspots <- paste(missD$viable_bloodspots, '(', missD$viable_bloodspots_perc, '%', ')')

missD$infant_dob <- as.character(missD$infant_dob)
missD$infant_dob_perc <- as.character(missD$infant_dob_perc)
missD$infant_dob <- paste(missD$infant_dob, '(', missD$infant_dob_perc, '%', ')')

missD$infantprophylaxis <- as.character(missD$infantprophylaxis)
missD$infantprophylaxis_perc <- as.character(missD$infantprophylaxis_perc)
missD$infantprophylaxis <- paste(missD$infantprophylaxis, '(', missD$infantprophylaxis_perc, '%', ')')

missD$pcrtest_requested <- as.character(missD$pcrtest_requested)
missD$pcrtest_requested_perc <- as.character(missD$pcrtest_requested_perc)
missD$pcrtest_requested <- paste(missD$pcrtest_requested, '(', missD$pcrtest_requested_perc, '%', ')')

missD$mother_pmtctregimen <- as.character(missD$mother_pmtctregimen)
missD$mother_pmtctregimen_perc <- as.character(missD$mother_pmtctregimen_perc)
missD$mother_pmtctregimen <- paste(missD$mother_pmtctregimen, '(', missD$mother_pmtctregimen_perc, '%', ')')

missD$pcrtestpositiveresult_date <- as.character(missD$pcrtestpositiveresult_date)
missD$pcrtestpositiveresult_date_perc <- as.character(missD$pcrtestpositiveresult_date_perc)
missD$pcrtestpositiveresult_date <- paste(missD$pcrtestpositiveresult_date, '(', missD$pcrtestpositiveresult_date_perc, '%', ')')

missD %>%
  filter(lab_name %in% c('AMPATH, Eldoret', 'Coast Provincial General Hospital', 'KEMRI CDC HIV/R, Kisumu', 'KEMRI CIPDCR, Alupe Busia', 'National HIV Reference Laboratory', 'KEMRI CVR HIV-P3, Nairobi')) %>%
  select(lab_name, sid, date_dbscollected, viable_bloodspots, infant_dob, infantprophylaxis, mother_pmtctregimen, pcrtest_requested, pcrtestpositiveresult_date) %>%
  rename('Lab Name'=lab_name, 'SID - N(%)'=sid, 'DBS collection date - N(%)'=date_dbscollected, 'viable blood spots - N(%)'=viable_bloodspots, 'infant DoB - N(%)'=infant_dob, 'infant prophylaxis - N(%)'=infantprophylaxis, 'mother PMTCT regimen - N(%)'=mother_pmtctregimen, 'requested PCR Test - N(%)'= pcrtest_requested, 'PCR test positive result date - N(%)'=pcrtestpositiveresult_date) %>%
  regulartable() %>%
  theme_box()
 
```
&nbsp;


# **4. Inconsistent Data:**
&nbsp;

  a. ## Inconsistent SID format

```{r wrongSID}
subset(data, substr(data$sid, 10,12) != grep('[[:digit:]]{3}$', substr(data$sid, 10, 12), value = T) | 
         substr(data$sid, 8,9) != grep('[[:digit:]]{2}$', substr(data$sid, 8, 9), value = T) |
         substr(data$sid, 1,2) != grep('[[:digit:]]{2}$', substr(data$sid, 1, 2), value = T) |
         substr(data$sid, 3,7) !='EIDKE' | (sapply(data$sid,nchar)!=12), 
       select = c('lab_name','sid')) %>%
  regulartable() %>% 
  theme_zebra() %>% 
  autofit()


```
&nbsp;

  b. ## *Date sample added to database* EARLIER than *date DBS sample collected*

```{r sampleDateEarlier}
subset(data, date_sampleadded < date_dbscollected, select = c('lab_name', 'sid', 'date_dbscollected', 'date_sampleadded' ))%>%
  regulartable() %>% 
  theme_zebra() %>% 
  autofit()
```

  c. ## Date Sample DBS Collected is earlier than Infant DoB 

```{r}
x <- subset(data, date_dbscollected < infant_dob, 
            select = c('lab_name', 'sid','date_dbscollected', 'infant_dob')) 
x$infant_dob <-  as.character(format(x$infant_dob, '%d/%m/%Y')) 
x$date_dbscollected <-  as.character(format(x$date_dbscollected, '%d/%m/%Y')) 
  regulartable(x) %>%
  theme_zebra() %>%
  autofit()


```
 &nbsp;
 
 d. ## Infant Age is greater than 18 months [547 days] as at dbs collection date
```{r}
x <- subset(data, infantage_days > 547, select = c('lab_name', 'sid', 'date_dbscollected', 'infant_dob', 'infantage_days' )) 
x$infant_dob <-  as.character(format(x$infant_dob, '%d/%m/%Y')) 
x$date_dbscollected <-  as.character(format(x$date_dbscollected, '%d/%m/%Y')) 
x$infantage_days <- as.character(format(x$infantage_days, digits=1))
  regulartable(x) %>%
  theme_zebra() %>%
  autofit()
```
&nbsp;

  e. ## Duplicate SIDs

```{r dups}
subset(data, duplicated(data$sid), select = c('lab_name', 'sid'))%>%
  regulartable() %>% 
  theme_zebra() %>% 
  autofit()

```

```{r cleanup}
rm(list = ls())

```


