---
title: "EID Updates"
author: "Daniel Fedha"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  word_document: 
    highlight: tango
    toc: yes
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, include = T, warning = F, message = F, comment = NA, collapse = T)
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "C:\\Program Files\\MiKTeX 2.9\\miktex\\bin\\x64", sep=.Platform$path.sep))
```
######
######
######
######
######
&nbsp;



```{r dbCon, echo=FALSE, include=FALSE}
#install.packages("RPostgreSQL", dependencies = T)
# # require("RPostgreSQL")
# 
# # 
# # # create a connection
# # pw <- {
# #   "0k0@us3r"
# # }
# 
# # 
# # # load Postgres driver
# # drv <- dbDriver("PostgreSQL")
# # 
# # # create a connection to okoa db
# # con <- dbConnect(drv, dbname = "okoa",
# #                  host = "localhost", port = 5432,
# #                  user = "okoauser", password = pw)
# # rm(pw)
# # # check for a table
# # # dbExistsTable(con, "okoa_laboratorysample")
# # 
# 
# # data <- dbGetQuery(con,
# # "SELECT	a.study_identification_number as sid,
#     #  		a.user_entering_sample_details_id as staff_id,
#     # 		g.name as lab_name,
#     #  		a.dbs_collection_date as date_dbscollected,
#     # 		a.viable_blood_spots_count as viable_bloodspots,
#     # 		a.date_of_birth as infant_dob,
#     # 		a.date_sample_details_added as date_sampleadded,
#     # 		a.infant_gender,
#     # 		c.name as sample_entrypoint,
#     # 		a.sample_entry_point_other as sample_entrypoint_other,
#     # 		e.name as infantprophylaxis,
#     #     d.name as infant_feedinginfo,
#     # 		a.infant_prophylaxis_regimen_other as infantprophylaxis_other,
#     # 		DATE_PART('day',dbs_collection_date::timestamp - date_of_birth::timestamp) as infantage_days,
#     # 		a.mother_age,
#     # 		a.mother_alive_status as mother_status,
#     # 		a.maternal_haart_initiation_date as motherhaartinitiation_date,
#     # 		b.name as  mother_pmtctregimen,
#     # 		a.maternal_pmtct_regimen_other as mother_pmtctregimen_other,
#     # 		f.name as pcrtest_requested,
#     # 		a.sample_redraw_pcr_code_id as sampleredraw_pcr,
#     # 		a.user_receiving_sample_dr_lab_id as staff_receivingsample,
#     # 		a.dr_lab_no_viable_blood_spots  as viablebloodspots_dr,
#     # 		a.date_genotyping_dbs_spots as dbsgenotyping_date,
#     # 		a.pcr_test_positive_result_date as pcrtestpositiveresult_date,
#     # 		a.dbs_placed_in_freezer_date as dbsplacedinfreezer_date,
#     # 		a.success_sequencing_rt_pol1_region,
#     # 		a.success_sequencing_pr_pol1_region
# # FROM
# # 		    public.okoa_laboratorysample as a
# # 
#         # left join okoa_motherpmtctregimen as b
#         # 	on a.maternal_pmtct_regimen_id = b.id
#         # 
#         # left join okoa_sampleentrypoint as c
#         # 	on a.sample_entry_point_id = c.id
#         # 
#         # left join okoa_infantfeedinginformation as d
#         # 	on a.infant_feeding_information_id = d.id
#         # 
#         # left join okoa_infantprophylaxisregimen as e
#         # 	on a.infant_prophylaxis_regimen_id = e.id
#         # 
#         # left join okoa_pcrtest as f
#         # 	on a.requested_pcr_test_id = f.id
#         # 
#         # left join okoa_laboratory as g
#         # 	on a.laboratory_id = g.id;")
#        
# 
# # #close db connection
# # lapply(dbListConnections(drv = dbDriver("PostgreSQL")), function(x) {dbDisconnect(conn = x)})
# 
# # data$enrolled = 1# 
# # saveRDS(data, file='C:\\Users\\Daniel Fedha\\Box Sync\\data\\EID\\data.rds')

```


```{r echo=FALSE}
library(ggplot2)
library(readr)
library(dplyr)
library(knitr)
library(flextable)
library(tidyr)

data <- readRDS('C:\\Users\\Daniel Fedha\\Box Sync\\data\\EID\\data.rds')
data$date_dbscollected <- as.Date(data$date_dbscollected, format = '%d/%m/%Y')
data$date_sampleadded <- as.Date(data$date_sampleadded, format = '%d/%m/%Y')
data$dbsplacedinfreezer_date<- as.Date(data$dbsplacedinfreezer_date, format = '%d/%m/%Y')
data$dbsgenotyping_date <- as.Date(data$dbsgenotyping_date, format = '%d/%m/%Y')
data$pcrtestpositiveresult_date <- as.Date(data$pcrtestpositiveresult_date, format = '%d/%m/%Y')
data$infant_dob <- as.Date(data$infant_dob, format = '%d/%m/%Y')
data$lab_name <- as.factor(data$lab_name)
data$infant_gender <- as.factor(data$infant_gender)
data$sample_entrypoint <- as.factor(data$sample_entrypoint)
data$infantprophylaxis <- as.factor(data$infantprophylaxis)
data$infant_feedinginfo <- as.factor(data$infant_feedinginfo)
data$mother_status <- as.factor(data$mother_status)
data$mother_pmtctregimen <- as.factor(data$mother_pmtctregimen)
data$pcrtest_requested <- as.factor(data$pcrtest_requested)
data$week_enrol <- as.integer(strftime(data$date_sampleadded, format = '%V'))-22

my_theme <- function(){
  theme_light() +
    theme(panel.grid = element_blank(),
          #panel.border = element_blank(),
          # axis.title = element_blank(),
          # axis.ticks = element_blank(),
          # axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = 8, colour = 'black'))
  }

```

# 1. **Enrolment Statistics**

a. ## Weekly Enrolment by Lab - Table:

```{r enrolLab, fig.height=2, fig.width=6}
sampleSize <- read_csv(file = 'C:/Users/Daniel Fedha/Box Sync/data/EID/sample_allocation.csv')
dtm <- merge(
  data %>%
    group_by(lab_name) %>%
    select(lab_name, enrolled, week_enrol) %>% 
    filter(week_enrol==1) %>%
    summarise('week 1'=sum(enrolled)),
  data %>%
    group_by(lab_name) %>%
    select(lab_name, enrolled, week_enrol) %>% 
    filter(week_enrol==2) %>%
    summarise('week 2'=sum(enrolled)),
        by='lab_name', all = T) %>%
  merge(data %>%
    group_by(lab_name) %>%
    select(lab_name, enrolled, week_enrol) %>% 
    filter(week_enrol==3) %>%
    summarise('week 3'=sum(enrolled)), 
  by='lab_name', all = T) %>% 
  merge(data %>%
          group_by(lab_name) %>%
          select(lab_name, enrolled, week_enrol) %>% 
          filter(week_enrol==4) %>%
          summarise('week 4'=sum(enrolled)),
        by='lab_name', all = T) %>%
  replace_na(list('week 1'=0, 
                  'week 2'=0,
                  'week 3'=0,
                  'week 4'=0))

dtm <- merge(dtm, sampleSize, by='lab_name') %>%
  rename('Lab Name'=lab_name)

dtm$`Lab Name` <- as.character(dtm$`Lab Name`)
dtm <- rbind(dtm, c("KEMRI CVR HIV-P3, Nairobi",0,0,0,0,125,0,0))
dtm <- rbind(dtm, c("Kenyatta National Hospital",0,0,0,0,23,0,0))
dtm <- rbind(dtm, c("KEMRI/WRP CRC, Kericho",0,0,0,0,65,0,0))

dtm$`week 1`<- as.integer(dtm$`week 1`)
dtm$`week 2`<- as.integer(dtm$`week 2`)
dtm$`week 3`<- as.integer(dtm$`week 3`)
dtm$`week 4`<- as.integer(dtm$`week 4`)

dtm$allocated_sampleSize <- as.integer(dtm$allocated_sampleSize)
dtm$Total <- dtm$`week 1`+ dtm$`week 2`+ dtm$`week 3`+ dtm$`week 4`
dtm$difference <- dtm$Total - dtm$allocated_sampleSize
dtm$percentage_achieved <- (dtm$Total/dtm$allocated_sampleSize)*100
dtm$percentage_achieved <- as.integer(format(dtm$percentage_achieved, digits = 2)) 
dtm <- dtm %>% arrange(desc(percentage_achieved))
dtm <- rbind(dtm, c("Total", colSums(dtm[,2:9])))

dtm$percentage_achieved[dtm$allocated_sampleSize==560] <-format((sum(as.integer(dtm$Total[1:8]))/560)*100, digits = 2)

dtm %>% select('Lab Name', `week 1`, `week 2`, `week 3`, `week 4`, Total, allocated_sampleSize, difference, percentage_achieved) %>%
  rename('% Achieved'= percentage_achieved, 'Sample Size'=allocated_sampleSize,
         'wk1'=`week 1`,'wk2'=`week 2`,'wk3'=`week 3`,'wk4'=`week 4`) %>%
  regulartable() %>%
  theme_zebra() %>%
  autofit()

```
Summary:&nbsp;
```{r summary}
sampEP <- aggregate(data$enrolled, by=list(sEP=data$sample_entrypoint), FUN=sum)
infaF <- aggregate(data$enrolled, by=list(infaF=data$infant_feedinginfo), FUN=sum)
pcrT <- aggregate(data$enrolled, by=list(pcrT=data$pcrtest_requested), FUN=sum)
pmtct <- aggregate(data$enrolled, by=list(infaF=data$mother_pmtctregimen), FUN=sum)

```

* `r sum(as.integer(dtm$Total[1:8]))` samples have been enrolled over a period of 4 weeks &nbsp;

  * `r subset(dtm, dtm$Total==max(as.integer(dtm[,'Total'][1:8])), select='Lab Name')` has enrolled the highest no. of samples: `r max(as.integer(dtm[,'Total'][1:8]))`, &nbsp;
  
  * 3 labs have not enrolled any sample:&nbsp;
  
    * KNH, P3, and NHRL. WRP is inactive &nbsp;
    
  * most samples [N=`r max(sampEP[,'x']) `, 
  (`r format(((max(sampEP[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)`%)] are through MCH/PMTCT sample entry point &nbsp;
  
  * most samples [N= `r max(infaF[,'x']) `,
  (`r format(((max(infaF[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)`%] are on BF &nbsp;
  
  * Initial PCR at 6 weeks or first contact is the most requested [N= `r max(pcrT[,'x']) `, 
  (`r format(((max(pcrT[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)`%] &nbsp;
  
  * most mothers [N= `r max(pmtct[,'x']) `, (`r format(((max(pmtct[,'x']))/sum(as.integer(dtm$Total[1:8])))*100,digits=2)` %] are on TDF+3TC+EFV &nbsp;
  
  * Missing data for all primary variables is 0% &nbsp;
  
  * No incosistency in SIDs nor dates.
&nbsp;

b. ##  Lab Enrolment by Gender - Bar Graph

```{r enrolBar, fig.height=2, fig.width=6}
data %>%
    group_by(lab_name, infant_gender) %>%
    select(lab_name, enrolled, infant_gender) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(lab_name, enrolled),
             y=enrolled,
             fill=infant_gender,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0)) +  # Set the axes to cross at 0
  labs(x='Laboratory',
       y='No. Enrolled') +
  my_theme() 

```

 &nbsp;
 
 c. ## Sample Entry Point by Lab - Bar Graph:

```{r enrolBSampEntryP, fig.height=3, fig.width=7}
data %>%
    group_by(sample_entrypoint, lab_name) %>%
    select(lab_name, enrolled, sample_entrypoint) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(sample_entrypoint, enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 50, 5)) +  # Set the axes to cross at 0
  
  labs(x='Laboratory',
       y=' ') +
  my_theme() 

```
&nbsp;

d. ## Infant Feeding by Lab - Bar Graph:

```{r fig.height=3, fig.width=7}
data %>%
    group_by(infant_feedinginfo, lab_name) %>%
    select(lab_name, enrolled, infant_feedinginfo) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(infant_feedinginfo,  enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 50, 5)) +  # Set the axes to cross at 0
  labs(x='Laboratory',
       y=' ') +
  my_theme() 

```

&nbsp;

  e. ## PCR Test Requested by Lab - Bar Graph:

```{r fig.height=2, fig.width=8}
data %>%
    group_by(pcrtest_requested, lab_name) %>%
    select(lab_name, enrolled, pcrtest_requested) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(pcrtest_requested, enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 50, 5)) +  # Set the axes to cross at 0
  labs(x='Laboratory',
       y=' ') +
  my_theme() 

```
&nbsp;

  f. ## Mother PMTCT Regimen by Lab - Bar Graph:
```{r fig.height=4, fig.width=7}

data %>%
    group_by(mother_pmtctregimen, lab_name) %>%
    select(lab_name, enrolled, mother_pmtctregimen) %>% 
    summarise('enrolled'=sum(enrolled)) %>%
  ggplot(aes(x=reorder(mother_pmtctregimen, enrolled),
             y=enrolled,
             fill=lab_name,
             position='stack')) +
  geom_bar(width=0.7, stat = 'identity') +
  geom_text(aes(label = enrolled,
                y=enrolled), 
            position=position_stack(vjust = 0.5)) + 
  coord_flip() +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 40, 5)) +  # Set the axes to cross at 0
  labs(x='Laboratory',
       y=' ') +
  my_theme() 


```
&nbsp;

g. ## Age distribution - Histogram

```{r enrolHist, fig.height=2, fig.width=6}

ggplot(data, 
           aes(data$infantage_days)) +
      geom_histogram(breaks=seq(1, 547, by =30),
                     col="red", 
                     aes(fill=..count..)) +
      scale_fill_gradient("Count", 
                          low = "green", 
                          high = "red") +
      labs(x='Age in Days',
           y='Count')+
  scale_y_continuous(expand = c(0, 0)) +  # Set the axes to cross at 0
  scale_x_continuous(expand = c(0, 0)) +  # Set the axes to cross at 0
  my_theme()
```


# **2. Missing Data:**

N=total no. missing per variable, &nbsp;

% = percentage of missing data per variable

```{r missingSID}
 missD <- merge(
   data %>%
     group_by(lab_name) %>%
     select(sid) %>% 
     summarise('sid'=sum(is.na(.))),
   data %>%
     group_by(lab_name) %>%
     select(date_dbscollected) %>% 
     summarise('date_dbscollected'=sum(is.na(.))),
   by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(viable_bloodspots) %>% 
      summarise('viable_bloodspots'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(infant_dob) %>% 
      summarise('infant_dob'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(infantprophylaxis) %>%
      summarise('infantprophylaxis'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(mother_pmtctregimen) %>% 
      summarise('mother_pmtctregimen'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(pcrtest_requested) %>% 
      summarise('pcrtest_requested'=sum(is.na(.))),
    by='lab_name', all=T) %>%
  merge(
    data %>%
      group_by(lab_name) %>%
      select(pcrtestpositiveresult_date) %>%
      summarise('pcrtestpositiveresult_date'=sum(is.na(.))),
    by='lab_name', all=T)

dtm2 <- dtm %>%
  na.omit() %>%
  select('Lab Name', Total) %>%
  rename(lab_name='Lab Name')
missD <- merge(missD, dtm2, by='lab_name', all = T)
rm(dtm2)
missD$sid_perc <- (missD$sid/as.integer(missD$Total))*100
missD$date_dbscollected_perc <- (missD$date_dbscollected/as.integer(missD$Total))*100
missD$viable_bloodspots_perc <- (missD$viable_bloodspots/as.integer(missD$Total))*100
missD$infant_dob_perc <- (missD$infant_dob/as.integer(missD$Total))*100
missD$infantprophylaxis_perc <- (missD$infantprophylaxis/as.integer(missD$Total))*100
missD$mother_pmtctregimen_perc <- (missD$mother_pmtctregimen/as.integer(missD$Total))*100
missD$pcrtest_requested_perc <- (missD$pcrtest_requested/as.integer(missD$Total))*100
missD$pcrtestpositiveresult_date_perc <- (missD$pcrtestpositiveresult_date/as.integer(missD$Total))*100

missD$sid <- as.character(missD$sid)
missD$sid_perc <- as.character(missD$sid_perc)
missD$sid <- paste(missD$sid, '(', missD$sid_perc, '%', ')')

missD$date_dbscollected <- as.character(missD$date_dbscollected)
missD$date_dbscollected_perc <- as.character(missD$date_dbscollected_perc)
missD$date_dbscollected <- paste(missD$date_dbscollected, '(', missD$date_dbscollected_perc, '%', ')')

missD$viable_bloodspots <- as.character(missD$viable_bloodspots)
missD$viable_bloodspots_perc <- as.character(missD$viable_bloodspots_perc)
missD$viable_bloodspots <- paste(missD$viable_bloodspots, '(', missD$viable_bloodspots_perc, '%', ')')

missD$infant_dob <- as.character(missD$infant_dob)
missD$infant_dob_perc <- as.character(missD$infant_dob_perc)
missD$infant_dob <- paste(missD$infant_dob, '(', missD$infant_dob_perc, '%', ')')

missD$infantprophylaxis <- as.character(missD$infantprophylaxis)
missD$infantprophylaxis_perc <- as.character(missD$infantprophylaxis_perc)
missD$infantprophylaxis <- paste(missD$infantprophylaxis, '(', missD$infantprophylaxis_perc, '%', ')')

missD$pcrtest_requested <- as.character(missD$pcrtest_requested)
missD$pcrtest_requested_perc <- as.character(missD$pcrtest_requested_perc)
missD$pcrtest_requested <- paste(missD$pcrtest_requested, '(', missD$pcrtest_requested_perc, '%', ')')

missD$mother_pmtctregimen <- as.character(missD$mother_pmtctregimen)
missD$mother_pmtctregimen_perc <- as.character(missD$mother_pmtctregimen_perc)
missD$mother_pmtctregimen <- paste(missD$mother_pmtctregimen, '(', missD$mother_pmtctregimen_perc, '%', ')')

missD$pcrtestpositiveresult_date <- as.character(missD$pcrtestpositiveresult_date)
missD$pcrtestpositiveresult_date_perc <- as.character(missD$pcrtestpositiveresult_date_perc)
missD$pcrtestpositiveresult_date <- paste(missD$pcrtestpositiveresult_date, '(', missD$pcrtestpositiveresult_date_perc, '%', ')')

missD %>%
  filter(lab_name %in% c('AMPATH, Eldoret', 'Coast Provincial General Hospital', 'KEMRI CDC HIV/R, Kisumu', 'KEMRI CIPDCR, Alupe Busia', 'National HIV Reference Laboratory')) %>%
  select(lab_name, sid, date_dbscollected, viable_bloodspots, infant_dob, infantprophylaxis, mother_pmtctregimen, pcrtest_requested, pcrtestpositiveresult_date) %>%
  rename('Lab Name'=lab_name, 'SID - N(%)'=sid, 'DBS collection date - N(%)'=date_dbscollected, 'viable blood spots - N(%)'=viable_bloodspots, 'infant DoB - N(%)'=infant_dob, 'infant prophylaxis - N(%)'=infantprophylaxis, 'mother PMTCT regimen - N(%)'=mother_pmtctregimen, 'requested PCR Test - N(%)'= pcrtest_requested, 'PCR test positive result date - N(%)'=pcrtestpositiveresult_date) %>%
  regulartable() %>%
  theme_box()
 
```
&nbsp;


# **3. Inconsistent Data:**
&nbsp;

  a. ## Wrong SID format

```{r wrongSID}
subset(data, substr(data$sid, 10,12) != grep('[[:digit:]]{3}$', substr(data$sid, 10, 12), value = T) | 
         substr(data$sid, 8,9) != grep('[[:digit:]]{2}$', substr(data$sid, 8, 9), value = T) |
         substr(data$sid, 1,2) != grep('[[:digit:]]{2}$', substr(data$sid, 1, 2), value = T) |
         substr(data$sid, 3,7) !='EIDKE' | (sapply(data$sid,nchar)!=12), 
       select = c('lab_name','sid')) %>%
  regulartable() %>% 
  theme_zebra() %>% 
  autofit()


```

  b. ## Date sample added to database earlier than Sample DBS collection date

```{r sampleDateEarlier}
subset(data, date_sampleadded < date_dbscollected, select = c('lab_name', 'sid', 'date_dbscollected', 'date_sampleadded' ))%>%
  regulartable() %>% 
  theme_zebra() %>% 
  autofit()
```

  c. ## Date Sample DBS Collected is earlier than Infant DoB or Infant Age is greater than 18 months

```{r}
x <- subset(data, date_dbscollected < infant_dob  | 
         infantage_days > 547, select = c('lab_name', 'sid', 'infant_dob', 'infantage_days' )) 
x$infant_dob <-  as.character(format(x$infant_dob, '%d/%m/%Y')) 
x$infantage_days <- as.character(format(x$infantage_days, digits=1))
  regulartable(x) %>%
  theme_zebra() %>%
  autofit()


```

  e. ## Duplicate SIDs

```{r dups}
subset(data, duplicated(data$sid), select = c('lab_name', 'sid'))%>%
  regulartable() %>% 
  theme_zebra() %>% 
  autofit()

```

